---
title: "Final project 0.3"
author: "Karam Abdin"
date: "2/16/2022"
output: html_document
---
Data description:
The data used in this visualization is from tidytuesday. The data dictionary could be find following this link: https://github.com/rfordatascience/tidytuesday/blob/master/data/2020/2020-07-07/readme.md#coffee_ratingscsv.

In this visualization I am focusing on:

1) The move of flavor and aroma colored with species.

2) Where the coffee come from defining by species using map graph.

3) Explore the country cups rating by scatter plot and show the quantities for each country in average.

4) Plotting for each coffee species how many rating according to the expiration years.

5) The change of bags tested over year in different bean color.

6) Crating the defects category for the elements (category_one_defects, category_two_defects) by the year.



```{r setup, include=FALSE}

library(tidyverse)
library(data.table)
library(openxlsx)
library(kableExtra)
library(modelsummary)
library(maps)
library(countrycode)
library(pacman)
library(ggplot2)
library(data.table)
library(rnaturalearth)
library(gganimate)
library(gifski)
library(ggthemes)
library(ggiraph)

```

```{r echo=TRUE, message=FALSE, warning=FALSE}
## Download Official repo for the tidytuesday project.
coffee_ratings <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-07-07/coffee_ratings.csv')

## Confirm the database to data table.
coffee_ratings <- data.table(coffee_ratings)
coffee_ratings <- coffee_ratings[,c("owner","farm_name","lot_number","mill","ico_number", "company","altitude",
                                    "region", "bag_weight","in_country_partner","harvest_year","owner_1","variety",
                                    "processing_method","moisture", "quakers",
                                    "certification_body","certification_address","certification_contact",
                                    "unit_of_measurement","altitude_low_meters",
                                    "altitude_high_meters","altitude_mean_meters" ):=NULL]

## Delete the unnecessary samples from country of origin column.
coffee_ratings$country_of_origin<- gsub("Cs*\\(.*", "",  coffee_ratings$country_of_origin)

## Finding the country code. 
coffee_ratings$country_code <- countrycode(coffee_ratings$country_of_origin, origin = "country.name",
                                           destination = "iso3c")

## Find the number of countries and the average of rating by country of origin.
Avr_rating <- coffee_ratings[,.(count = .N, mean_country_rating = mean(total_cup_points),
                         adm0_a3 = country_code, species = species), by = country_of_origin]
#str(coffee_ratings)

coffee_ratings_head <- head(coffee_ratings)
kable(coffee_ratings_head,  caption = " This table presents the dataset's head of the coffee ratings")

```
```{r echo=TRUE, warning=FALSE}
## Create the year column from grading date
coffee_ratings$year = substring(gsub(".*,","",coffee_ratings$grading_date),2,5)

producer_rank <- coffee_ratings[!is.na(producer), .(.N,
                                    avg_aroma =  round(mean(aroma),digits = 2),
                                    avg_flavor = round(mean(flavor),digits = 2)),
                                by = .(year, species)]

reg_1 <- lm( avg_aroma ~ avg_flavor , data = producer_rank )

msummary(list(reg_1),
         fmt="%.4f",
         gof_omit = 'DF|Deviance|Log.Lik.|F|R2 Adj.|AIC|BIC',
         stars=c('*' = .05, '**' = .01))
```


```{r echo=TRUE, warning=FALSE}
# Crate a theme function
theme_custom <- function() {
  theme(
    axis.text = element_text(
      family = 'Times New Roman',
      color  = "#e6550d",
      size   = 13,
      face   = "bold"),
    axis.title = element_text(
      family = 'Times New Roman',
      color  = "#636363",
      size   = 13,
      face   = "bold"),
    axis.text.y = element_text(angle = 90, hjust = 0.5),
    panel.background = element_rect(
      fill = "#f0f0f0",
      color = "#f0f0f0",
      size = 2)
  )
}
```


1) The move of flavor and aroma colored with species.

```{r echo=TRUE, warning=FALSE}


## Crate an animation point to show the move of flavor and aroma colored with species.
graph1 <- ggplot(producer_rank, aes(avg_aroma, avg_flavor, color = species )) +
  geom_abline(slope = 1, intercept = 0) +
  geom_point(size = 6) +
  theme_bw() +
  theme(legend.position = "top")+
  transition_time(as.integer(year)) +
  shadow_wake(wake_length = .1, alpha = FALSE) +
  labs(title = "year: {frame_time}",
       x = "Aroma",
       y = "Favor")

graph1 + theme_custom()
```

This point graph presents the Favor grade & Aroma grade for each year. From the move of the point we can conclude that the to grades do not have any relationship in common. Where we can see from the regression table this two grades are statistical Significant.  

\n---------------------------------------------------------------------------------------------
\n---------------------------------------------------------------------------------------------


2) Where the coffee come from defining by species using map graph.
```{r echo=TRUE, warning=FALSE}
## Delete the unnecessary samples from country of origin column.
coffee_ratings$country_of_origin<- gsub("Cs*\\(.*", "",  coffee_ratings$country_of_origin)

## Finding the country code. 
coffee_ratings$country_code <- countrycode(coffee_ratings$country_of_origin, origin = "country.name",
                                           destination = "iso3c")
## Find the number of countries and the average of rating by country of origin.
Avr_rating <- coffee_ratings[,.(count = .N, mean_country_rating = mean(total_cup_points),
                         adm0_a3 = country_code, species = species), by = country_of_origin]
## Creating the data map.
world <- ne_countries(scale = "medium", returnclass = "sf")

## Merge the data files.
Final_coffee_ratting <- merge(world, Avr_rating, by= "adm0_a3")

## Mapping the data using geom_sf by species.
ggplot(data = Final_coffee_ratting) +
  borders() +
  geom_sf(aes(fill = mean_country_rating, geometry = geometry)) +
  scale_fill_viridis_c(trans = "sqrt") +
  theme(legend.position = "bottom") +
  coord_sf(ylim = c(-50, 80)) +
  facet_wrap(~ species, nrow = 2) +
  labs(title = "Average Coffee Rating:",
       subtitle = "By country and species",
       fill = "Average Rating / 100",
       x = "",
       y = "")
```

\n The map graph shows the countries that produce the coffee-related to the species where it is clear that is not all the countries which produce arabica coffee they produce Robusta coffee.
 
\n---------------------------------------------------------------------------------------------
\n---------------------------------------------------------------------------------------------

 
3) Explore the country cups rating by scatter plot and show the quantities for each country in average.
```{r echo=TRUE, warning=FALSE}

## Create the company cups columns (average of total cups) by species and country code.
company_cups <- coffee_ratings[, .(.N,total_cups = round(mean(total_cup_points, na.rm = TRUE), digits = 2)),
                               by =.(species, country_code)]

## The information that included in tooltip.
text <- paste("Country:", company_cups$country_code,
              "<br>Total cups on average: ", company_cups$total_cups)

## Crate point interactive graph using girafe.
company_plot <- ggplot(company_cups, aes(total_cups,
                                         N,
                                         color = species,
                                         tooltip = text )) +
  geom_point_interactive() +
  scale_color_brewer(type = 'qual', palette = 'Dark2', direction = 10) +
  theme_gdocs() +
  scale_fill_gdocs() +
  labs(title = "Average Coffee Rating:",
       subtitle = "By country and species",
       fill = "Average Rating / 100",
       caption = "Please: Select a point for more details",
       x = "",
       y = "")

girafe(ggobj = company_plot)

```
This point graph presents all the countries that produce coffee divided by species. In this graph, we can easily select a point to see which country it presents and details about the total cups on average that the country produces.

\n---------------------------------------------------------------------------------------------
\n---------------------------------------------------------------------------------------------



4) Plotting for each coffee species how many rating according to the expiration years.
```{r echo=TRUE, warning=FALSE}

## Cleaning the expiration column to just return the year's expiration.
experation_year <- coffee_ratings[,.(.N, species = species, year = gsub(".*,","",coffee_ratings$expiration))]
## Counting the expiration year for each year.
experation_year<- experation_year %>% select(species, year) %>% count(year, species)

exp_plot <- ggplot(data = experation_year, aes( x = species, y = n , fill= year)) +
  geom_col() +
  theme_stata() + scale_fill_stata() +
  theme( legend.position = "none") +
  transition_time(as.integer(year)) +
  labs(title =  "Average Coffee Rating in year( {frame_time} )",
       subtitle = "By species",
       x = "",
       y = "")
  

animate(exp_plot, renderer = gifski_renderer())

```

\n This bar graph presents the different between the quantity of producing Arabica and Robusta coffee in each year relating to our data. It shows that the Arabica coffee produces almost 80% in average more than Robusta coffee.

\n---------------------------------------------------------------------------------------------
\n---------------------------------------------------------------------------------------------


5) The change of bags tested over year in different bean color.
```{r echo=TRUE, warning=FALSE}
 

## The change of bags tested over year in different bean color.
## The colors of coffee with the count of bags numbers by color & year.
coffee_ratings$year = substring(gsub(".*,","",coffee_ratings$grading_date),2,5)

coffee_color <- coffee_ratings[!color == "None"][!is.na(color), .(number = .N, Number_of_bags = sum(number_of_bags)),
                                                 by = .(Color_of_bean = color, year)]
## The change of the bags numbers accourding to the years 
Color <- ggplot(coffee_color, aes(x = year, y = Number_of_bags, color = Color_of_bean)) +
  stat_summary(fun = sum,geom='line',aes(group = Color_of_bean,
                                         color = Color_of_bean), size = 1.3) +
  theme_base() +
  theme(legend.position = "bottom") + 
  xlab("Year") + ylab("Number of bags tested") +
  labs(subtitle = "Number of bags tested over year with different bean color") +
  transition_reveal(as.integer(year))
  
## Plot the graph in animation mode.
animate(Color, renderer = gifski_renderer())
```

This line graph shows the difference between the color of beans. From the graph, we can say that the Green bean is the most used bean, especially in 2014,2015 years compared with Blue-green and Bluish-green. Where the other two beans move closely in this timeline.

\n---------------------------------------------------------------------------------------------
\n---------------------------------------------------------------------------------------------


6) Crating the defects category for the elements (category_one_defects, category_two_defects) by the year.
```{r echo=TRUE, warning=FALSE}

## Create the year column from grading date
coffee_ratings$year = substring(gsub(".*,","",coffee_ratings$grading_date),2,5)

## Creating a category table counts the average of one & two category defects by the year.
categorys <- coffee_ratings[,.(.N, category_one_defects = round(mean(category_one_defects), digits = 2),
                               category_two_defects = round(mean(category_two_defects),digits = 2)), by = year]

## Create two lines for each category defect on one plot which shows the different between them.
category_plot <- ggplot( categorys, aes(x=year,  group = 1, tooltip = "text")) + 
  geom_line(aes(y=as.character(category_two_defects), col="category_two_defects"),size = 1.5) + 
  geom_line(aes(y=as.character(category_one_defects), col="category_one_defects"),size = 1.5) + 
  theme_bw() +
  theme(legend.position = "bottom") +
  scale_color_brewer(type = 'seq', palette = 'Reds', direction = 10) +
  labs(title="Category Defects", 
       y="Category Defects",
       x="Year")

girafe(ggobj = category_plot)

```
The dark orang line presents the coffee with two defects and the light color represents the coffee with one defect. From the graph, It is clear that the coffee with two defects is higher than the coffee with one defect during the years but they move almost in the same direction. After 2016 we see that the coffee with one defect moved in the opposite direction. In this case, we have two assumptions:
  \n 1- We do not have enough data after 2016 about the one defect coffee.
  \n 2- The one defect of coffee has been improved.
  
\n---------------------------------------------------------------------------------------------
\n---------------------------------------------------------------------------------------------






